---
title: "The Effects of School Proximity and the School Social Index on Housing Prices" 
subtitle: "\\presentationdate"
author: 
  - name: Yanyi Ji
  - name: Daniel Tobien
  - name: Luc Wichtmann
format: 
  beamer:
    theme: "metropolis"
    colortheme: "seahorse"
    include-in-header: custom_theme.tex
bibliography: references_results.bib
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| include: false
source("Final Model mit Fragen.R")
```

## Agenda

**1. Motivation, Existing Literature & Research Question**\
*Why school proximity and social indices matter.*

\vspace{0.3em}

**2. Data & Methodology**\
*Data used and our empirical model.*

\vspace{0.3em}

**3. Results & Main Findings**\
*Presentation and interpretation of our findings.*

\vspace{0.3em}

**4. Limitations & Conclusion**\
*Limitations of our work and concluding thoughts.*

\vspace{0.3em}

# Motivation, Existing Literature & Research Question

## Motivation

-   **Personal Interest**: Analyzing the real estate market's reaction to various factors is particularly interesting from an economic standpoint, as it offers insights into market dynamics. Additionally, the topic is personally intriguing, adding an extra layer of motivation to explore it further.

-   **Academic and Practical Relevance**: Housing markets are influenced by various social and economic factors, and school quality is also cited as a underestimated determinant of property values, especially in research [@Seo_Simons_2009].

## Motivation

Understanding the relationship between school distance, school quality and property prices can provide valuable insights for researchers and individuals interested in the real estate market.

## Related Literature Background

Several papers have examined the effect of proximity to schools on housing prices:

-   **Empirical Strategies**: While hedonic models are the dominant approach to estimating housing price effects, alternative empirical strategies have also been proposed [see @Black_Machin_2011]

-   **Findings**: Empirical evidence suggests that proximity to schools positively affects housing prices [@Rosiers_Lagana_Theriault_2001; @Huang_Hess_2018]

## Related Literature Background

-   **Other Factors**: Related work also documents capitalization effects of school reputation [@Chin_Foong_2006], differences between public and private schools [@Sah_Conroy_Narwold_2016] and school quality [@Edusei_Espey_Lin_2007; @Metz_2015]

-   **Environment of studies**: Much of the literature focuses on settings with binding school catchment areas and measures school quality primarily through test scores rather than demographic composition.

## Research Question

We study a setting *without binding school catchment areas* (NRW, Germany) and assess whether both *school proximity* and a *demographically based school* *social index* are capitalized into housing prices.

## Research Question

**Research Question:** How do school proximity and school social indices affect the prices of detached houses, and do these effects differ between primary and secondary schools in North Rhine–Westphalia?

# Data & Methodology

## Data

-   **Housing Data**: Cross-section dataset for houses for sale published by the Research Data Center Ruhr (FDZ Ruhr) [@RWI_ImmobilienScout24_2023].
-   **School Data**: School social index dataset from North Rhine-Westphalias Ministry of Schools and Education [@MSB_NRW_Schulsozialindex].

\footnotetext{Data used from both datasets refer to the year 2022.}

## Pre-Processing of Data

::: {.center data-latex=""}
```{mermaid}
%%| fig-height: 2.9

stateDiagram-v2
    direction LR

    Final: Final Dataset
    state Final {
        
        Housing: Houses for sale
        state Housing {
            [*] --> NA
            NA: Removing NA values
            NA --> File
            File: Including only single-family houses in NRW
            File --> Outl
            Outl: Removing outliers
            Outl --> Dedu
            Dedu: Eliminating duplicates 
            Dedu --> Muta
            Muta: Renaming columns
            Muta --> [*]
        }
        
        School_Dist: School Social Index and Distance
        state School_Dist {

            Index: School Social Index
            state Index {
                [*] --> Types
                Types: Grouping school types
                Types --> Remo
                Remo: Removing invalid values
                Remo --> Muta2
                Muta2: Renaming columns
                Muta2 --> [*]
            }


            Dist: Distance to Schools
            state Dist {
                [*] --> [*]
            }

        }

    }

```
:::

## Empirical Framework
We estimate regression-based hedonic models to 

- examine how school proximity and school characteristics are capitalized into 
housing prices and
- whether the impact of school proximity on housing prices differs by the 
school’s social index


## Empirical Framework: Hedonic Price Model
\begin{center}
\colorbox{blue!10}{
$\displaystyle
\log(P_i) = 
\beta_0 + 
\beta_1 \text{Dist}_i + 
X_i' \gamma + 
\varepsilon_i
$
}
\end{center}

\vspace{1.5em}

- $P_i$ denotes the listing price of house $i$
\vspace{0.1em}
- $\text{Dist}_i$ measures distance to the nearest school
\vspace{0.1em}
- $X_i$ is a vector of housing characteristics
\vspace{0.1em}
- $\gamma$ is a vector of coefficients corresponding to $X_i$
\vspace{0.1em}
- $\varepsilon_i$ captures unobserved factors

## Empirical Framework: Hedonic Price Model + Interaction

\begin{center}
\colorbox{blue!10}{
$\displaystyle
\log(P_i) = 
\beta_0 + 
\beta_1 ( \text{Index}_i \times \text{Dist}_i ) + 
X_i' \gamma
+ \varepsilon_i
$
}
\end{center}

\vspace{1.5em}

- $P_i$ denotes the listing price of house $i$
\vspace{0.1em}
- $\text{Index}_i$ is the school social index
\vspace{0.1em}
- $\text{Dist}_i$ measures distance to the nearest school
\vspace{0.1em}
- $X_i$ is a vector of housing characteristics
\vspace{0.1em}
- $\varepsilon_i$ captures unobserved factors

## Empirical Framework: Final Model


\begin{center}
\colorbox{blue!10}{
$\displaystyle
\begin{aligned}
\log(P_i)
&=
\beta_0
+ \beta_1 \text{Dist}_i
+ \beta_2 \text{Index}_i \\
&\quad
+ \beta_3 \left( \text{Dist}_i \times \text{Index}_i \right)
+ X_i' \gamma
+ \mu_{\ell(i)}
+ \varepsilon_i
\end{aligned}
$
}
\end{center}



- $P_i$ denotes the listing price of house $i$
\vspace{0.1em}
- $\text{Dist}_i$ measures distance to the nearest school
\vspace{0.1em}
- $X_i$ is a vector of housing characteristics
\vspace{0.1em}
- $\text{Index}_i$ is the school social index
\vspace{0.1em}
- $mu_{\ell(i)}$ are municipality fixed effects
\vspace{0.1em}
- $\varepsilon_i$ captures unobserved factors

# Results and Main Findings

## Visualization for Distance to Primary School

<!-- This part has to stay on the last page before the references to indicate
the end of the presentation for the progress bar defined in the custom 
theme file.
This does not work yet... -->
\setcontentend

```{r}
#| include: true 
Testplot
```

## Limitations
- **Direct Distance**: The distance is not measured from the houses to the nearest school directly 
but from the school to the center of the raster cell.
- **Fragmented Grid**: We do not have data for every grid cell of NRW but only a
fragmented set of grids containing houses from the available datasets.

## References {.allowframebreaks}

::: {#refs}
:::
