---
title: "The Effects of School Proximity and the School Social Index on Housing 
Prices"
subtitle: "\\presentationdate"
author: 
  - name: Yanyi Ji
  - name: Daniel Tobien
  - name: Luc Wichtmann
format: 
  beamer:
    theme: "metropolis"
    colortheme: "seahorse"
    include-in-header: custom_theme.tex
bibliography: references_results.bib

---

```{r}
#| echo: false
#| eval: false 
#| source: distance_analysis.R
```


## Agenda
**1. Motivation, Existing Literature \& Research Question**  
*Why school proximity and social indices matter.*

**2. Data \& Methodology**  
*Data used and our empirical model.*  

\vspace{0.1em}

**3. Results \& Main Findings**  
*Presentation and interpretation of our findings.*

\vspace{0.1em}

**4. Limitations \& Conclusion**  
*Limitations of our work and concluding thoughts.*

\vspace{0.1em}


# Motivation, Existing Literature \& Research Question
## Motivation
- **Personal Interest**: Analyzing the real estate market's reaction to various 
factors is particularly interesting from an economic standpoint, as it offers 
insights into market dynamics. Additionally, the topic is personally intriguing, 
adding an extra layer of motivation to explore it further.

- **Academic and Practical Relevance**: Housing markets are influenced by 
various social and  economic factors, and school quality is also cited as a 
underestimated determinant of property values, especially in research 
[@Seo_Simons_2009].


## Motivation
Understanding the relationship between school distance, school quality and 
property prices can provide valuable insights for researchers and individuals 
interested in the real estate market.  


## Related Literature Background
Several papers have examined the effect of proximity to schools on housing 
prices:

- **Empirical Strategies**: While hedonic models are the dominant approach to 
estimating housing price effects, alternative empirical strategies have also 
been proposed [see @Black_Machin_2011]

- **Findings**: Empirical evidence suggests that proximity to schools positively 
affects housing prices [@Rosiers_Lagana_Theriault_2001; @Huang_Hess_2018]


## Related Literature Background
- **Other Factors**: Related work also documents capitalization effects of school 
reputation [@Chin_Foong_2006], differences between public and private schools 
[@Sah_Conroy_Narwold_2016] and school quality [@Edusei_Espey_Lin_2007; 
@Metz_2015]

- **Environment of studies**: Much of the literature focuses on settings with 
binding school catchment areas and measures school quality primarily through 
test scores rather than demographic composition.


## Research Question
We study a setting *without binding school catchment areas* (NRW, Germany)
and assess whether both *school proximity* and a *demographically based school* 
*social index* are capitalized into housing prices.

## Research Question
**Research Question:** How do school proximity and school social indices affect 
the prices of detached houses, and do these effects differ between primary and 
secondary schools in North Rhine–Westphalia?


# Data \& Methodology
## Data
- **Housing Data**: Cross-section dataset for  houses for sale published by the 
Research Data Center Ruhr (FDZ Ruhr) [@RWI_ImmobilienScout24_2023].
- **School Data**: School social index dataset from North Rhine-Westphalias
Ministry of Schools and Education [@MSB_NRW_Schulsozialindex].

\footnotetext{Data used from both datasets refer to the year 2022.}


## Pre-Processing of Data
::: {.center  data-latex=""}
```{mermaid}
%%| fig-height: 2.9

stateDiagram-v2
    direction LR

    Final: Final Dataset
    state Final {
        
        Housing: Houses for sale
        state Housing {
            [*] --> NA
            NA: Removing NA values
            NA --> File
            File: Including only single-family houses in NRW
            File --> Outl
            Outl: Removing outliers
            Outl --> Dedu
            Dedu: Eliminating duplicates 
            Dedu --> Muta
            Muta: Renaming columns
            Muta --> [*]
        }
        
        School_Dist: School Social Index and Distance
        state School_Dist {

            Index: School Social Index
            state Index {
                [*] --> Types
                Types: Grouping school types
                Types --> Remo
                Remo: Removing invalid values
                Remo --> Muta2
                Muta2: Renaming columns
                Muta2 --> [*]
            }


            Dist: Distance to Schools
            state Dist {
                [*] --> [*]
            }

        }

    }

```
:::


## Empirical Framework


# Results and Main Findings

## Visualization for Distance to Primary School


```{r}
source("Final Model mit Fragen.R")

ggplot(df_final_with_social_5km, aes(x = dist_primary_km, y = log_price, color = social_index_factor)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Influence of Distance to Primary School on House Prices by Social Index Level",
       x = "Distance to Primary School (km)",
       y = "Log-House Price",
       color = "Social Index Level") +
  scale_color_manual(values = c("#FFB6C1", "#98C9E6", "#FFD700", "#D3D3D3", "#8FBC8F", "#D2B48C", "#A9A9A9", "#E6E6FA", "#F08080")) +
  theme_minimal()
```

## References {.allowframebreaks}
::: {#refs}
:::
