---
title: "Project Outline Presentation"
subtitle: "Housing Prices & School Quality in NRW"
author: "Daniel Jeffrey Tobien, Yanyi Ji, Luc Wichtmann"
date: "`r Sys.Date()`"
format: beamer
aspectratio: 169  
theme: "Frankfurt"  
colortheme: "seahorse"  
fontsize: 11pt  
header-includes:
  - \usepackage{xcolor}  # Ermöglicht die Verwendung von Farben
  - \usepackage{fontawesome5}  # Fügt Symbole hinzu
  - \setbeamercolor{block title}{fg=white,bg=blue}  
  - \setbeamercolor{block body}{fg=white,bg=lightgray} 
  - \usepackage{graphicx}  
  - \usepackage{amsmath}  
  - \setbeamertemplate{footline}{\hfill\insertframenumber\footnotesize{Data Analysis with R}}
editor: source  
---


## Agenda 


\frametitle{Agenda}

\begin{block}{Agenda}
  \begin{itemize}
    \item[] \faPlayCircle\ \textbf{Motivation}
    \item[] \faQuestionCircle\ \textbf{Research Question}
    \item[] \faDatabase\ \textbf{Data}
    \item[] \faCogs\ \textbf{Methods}
    \item[] \faArrowCircleRight\ \textbf{Next Steps}
  \end{itemize}
\end{block}

---

## Motivation  


•⁠  ⁠*Why is this interesting to us?*
  - Analyzing the real estate market's reaction to various factors is particularly interesting from an economic standpoint, as it offers insights into market dynamics. Additionally, the topic is personally intriguing, adding an extra layer of motivation to explore it further. 
  - We also believe that understanding the relationship between school quality and property prices can provide valuable insights for researchers and individuals interested in the real estate market.
  

•⁠  ⁠*Why is this important in a broader context?*
  - *Housing markets* are influenced by various social and economic factors, and school quality is also cited as a underestimated determinant of property values, especially in research.\footnote{Youngme Seo, Robert A. Simons, 2009}
  - Investigating how much the social index of schools influences housing prices, especially in relation to the proximity of schools, could offer an interesting perspective on the interplay between these factors in urban environments.

---

## Research Question 

\vspace{0.4cm}

\textbf{Research Question:}\
\textit{How does the social index of schools influence housing prices, and does the proximity to schools modulate this effect in urban areas?}

---

## Data 

•⁠  ⁠*Housing Data*(Campus File HK 2022):
  - Property prices (purchase price, price per sqm)
  - Characteristics: size, rooms, floor space, year built
  - Location (urban vs rural)

•⁠  ⁠*School Data*:
  - Social Index & quality of education
  - School types (public, private)
  - Proximity to schools (distance)

•⁠  ⁠*Regional Data*:
  - Demographic: income distribution, unemployment rates
  - Geographic classification (urban vs rural)
  
---
  
```{r}
# =======================================================
# 1. SETUP: Libraries & Global Theme
# =======================================================

library(tidyverse)
library(fs)
library(scales)   # For currency/number formatting
library(ggrepel)  # For non-overlapping labels
library(patchwork) # Optional: For combining plots

# Define a unified theme for all plots
theme_style <- theme_minimal(base_size = 14) +
  theme(
    text = element_text(family = "sans", color = "#333333"),
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5, margin = margin(b = 10)),
    plot.subtitle = element_text(color = "grey50", size = 12, hjust = 0.5, margin = margin(b = 20)),
    plot.caption = element_text(color = "grey60", size = 10, margin = margin(t = 15)),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(), # Cleaner look for histograms
    axis.text = element_text(color = "grey40"),
    axis.title = element_text(face = "bold", margin = margin(r = 10))
  )

# Set global default theme
theme_set(theme_style)

# =======================================================
# 2. DATA LOADING & CLEANING
# =======================================================

# --- A. Housing Data ---
housing_path <- "course_data/housing_data/cross_section/CampusFile_HK_2022.csv"

df_HK <- read_delim(housing_path, delim = ",", 
                    locale = locale(decimal_mark = "."), 
                    na = c("-5", "-6", "-7", "-8", "-9", "-11", "NA", ""),
                    show_col_types = FALSE) %>%
  filter(blid == "North Rhine-Westphalia") %>%
  mutate(filename = path_file(housing_path)) %>%
  # Robust number parsing
  mutate(grundstuecksflaeche_num = parse_number(as.character(grundstuecksflaeche)),
         wohnflaeche_num = parse_number(as.character(wohnflaeche)),
         kaufpreis_num = parse_number(as.character(kaufpreis))) %>%
  type_convert()

# --- B. School Data ---
school_path <- "course_data/school_data/2022_social_index.csv"

df_school <- read_delim(school_path, delim = ";", 
                        locale = locale(encoding = "Latin1", decimal_mark = ","),
                        show_col_types = FALSE)

# --- C. Distance Data ---
dist_path <- "course_data/school_data/distance_to_schools.csv"

df_dist <- read_delim(dist_path, delim = ",", 
                      locale = locale(decimal_mark = ","), 
                      show_col_types = FALSE)


# =======================================================
# 3. VISUALIZATIONS
# =======================================================

# -------------------------------------------------------
# Plot 1: Purchase Price (Raw Distribution)
# Goal: visualize the right-skewed market structure
# -------------------------------------------------------
p1_price_raw <- df_HK %>% 
  ggplot(aes(x = kaufpreis_num)) +
  geom_histogram(bins = 50, fill = "#2C3E50", color = "white", alpha = 0.9) +
  
  # Zoom in on 99% of data (hide extreme outliers without filtering them)
  coord_cartesian(xlim = c(0, quantile(df_HK$kaufpreis_num, 0.99, na.rm = TRUE))) +
  
  scale_x_continuous(labels = label_number(suffix = " €", scale = 1e-3, big.mark = ",")) +
  labs(title = "Distribution of Purchase Prices", 
       subtitle = "Zoomed in on main market (Top 1% outliers hidden)",
       x = "Price (in Thousand EUR)", 
       y = "Count")

# -------------------------------------------------------
# Plot 2: Purchase Price (Log Transformation)
# Goal: Check normality for regression models
# -------------------------------------------------------
p2_price_log <- df_HK %>% 
  filter(kaufpreis_num > 0) %>% # Log requires positive values
  ggplot(aes(x = log(kaufpreis_num))) +
  geom_histogram(aes(y = after_stat(density)), bins = 50, fill = "#2C3E50", color = "white", alpha = 0.5) +
  geom_density(color = "#E74C3C", linewidth = 1.2) +
  labs(title = "Log-transformed Price Distribution", 
       subtitle = "Approximates normal distribution",
       x = "Log(Price)", 
       y = "Density")

# -------------------------------------------------------
# Plot 3: Construction Year (Age Structure)
# Goal: Analyze building eras (focusing on post-1900)
# -------------------------------------------------------
p3_age <- df_HK %>%
  ggplot(aes(x = baujahr)) +
  geom_histogram(binwidth = 5, fill = "#2C3E50", color = "white", alpha = 0.9) +
  
  # Median Line
  geom_vline(aes(xintercept = median(baujahr, na.rm = TRUE)), 
             color = "#E74C3C", linetype = "dashed", linewidth = 0.8) +
  annotate("text", x = median(df_HK$baujahr, na.rm = TRUE) - 5, 
           y = 100, label = "Median Year", color = "#E74C3C", angle = 90, hjust = 0) +
  
  scale_x_continuous(limits = c(1900, 2025), breaks = seq(1900, 2025, by = 20)) +
  scale_y_continuous(labels = comma, expand = c(0, 0)) +
  labs(title = "Age Structure of Houses",
       subtitle = "Distribution of Construction Year (Baujahr)",
       x = "Construction Year",
       y = "Count")




# -------------------------------------------------------
# Plot 4: Plot Area (Log Scale Histogram)
# Goal: Visualize land size including small plots and large estates
# -------------------------------------------------------
p4_area_log <- df_HK %>%
  filter(!is.na(grundstuecksflaeche_num) & grundstuecksflaeche_num >= 50) %>% 
  ggplot(aes(x = grundstuecksflaeche_num)) +
  geom_histogram(bins = 50, fill = "#2C3E50", color = "white", alpha = 0.9) +
  
  # Log X-Axis
  scale_x_log10(breaks = c(50, 100, 200, 500, 1000, 5000, 10000, 50000),
                labels = comma_format(accuracy = 1),
                expand = c(0, 0),
                limits = c(50, NA)) +
  
  # Median Line
  geom_vline(aes(xintercept = median(grundstuecksflaeche_num, na.rm = TRUE)), 
             color = "#E74C3C", linetype = "dashed", linewidth = 1) +
  annotate("text", 
           x = median(df_HK$grundstuecksflaeche_num, na.rm = TRUE) * 1.5, 
           y = 3000, # Adjust based on max count
           label = paste("Median:", round(median(df_HK$grundstuecksflaeche_num, na.rm = TRUE))), 
           color = "#E74C3C", hjust = 0) +
  
  labs(title = "Distribution of Plot Area",
       subtitle = "Histogram with Logarithmic X-Axis (Starts from 50 sqm)",
       x = "Plot Area (sqm) - Log Scale", 
       y = "Count")




# -------------------------------------------------------
# Plot 5: Living Area vs. Price (Scatter)
# Goal: Sanity Check - Positive correlation expected
# -------------------------------------------------------
p5_sanity <- df_HK %>%
  filter(wohnflaeche_num > 10 & wohnflaeche_num < 600) %>%
  filter(kaufpreis_num > 10000 & kaufpreis_num < 3000000) %>%
  ggplot(aes(x = wohnflaeche_num, y = kaufpreis_num)) +
  
  # Small transparent points to handle overplotting
  geom_point(alpha = 0.1, color = "#2C3E50", size = 0.8) +
  geom_smooth(method = "lm", color = "#E74C3C", se = FALSE, linewidth = 1) +
  
  scale_y_continuous(labels = dollar_format(prefix = "€", big.mark = ",")) +
  scale_x_continuous(labels = comma_format(suffix = " m²")) +
  labs(title = "Sanity Check: Living Area vs. Price",
       subtitle = "Relationship between Wohnflaeche and Price",
       x = "Living Area", 
       y = "Price")




# -------------------------------------------------------
# Plot 6: Social Index Distribution
# Goal: Balance check of school locations
# -------------------------------------------------------
p6_social <- df_school %>%
  filter(!is.na(Sozialindexstufe)) %>%
  ggplot(aes(x = as.factor(Sozialindexstufe))) +
  geom_bar(fill = "#2C3E50", color = "white", alpha = 0.9) +
  labs(title = "Distribution of School Social Index",
       subtitle = "1 = Low Challenge, 9 = High Challenge",
       x = "Social Index Level", 
       y = "Number of Schools")




# -------------------------------------------------------
# Plot 7: Distance to Nearest School
# Goal: Verify walking distance (0.5 - 3km)
# -------------------------------------------------------
p7_distance <- df_dist %>%
  filter(nn_order == 1) %>% # Nearest school only
  mutate(dist_num = as.numeric(dist)) %>%
  filter(!is.na(dist_num) & dist_num < 15) %>% 
  
  ggplot(aes(x = dist_num)) +
  geom_histogram(binwidth = 0.2, fill = "#2C3E50", color = "white", alpha = 0.9) +
  
  # Highlight Walking Zone
  annotate("rect", xmin = 0.5, xmax = 3, ymin = 0, ymax = Inf, 
           fill = "#27AE60", alpha = 0.15) +
  annotate("text", x = 1.75, y = Inf, label = "Walking Zone (0.5-3km)", 
           color = "#27AE60", vjust = 1.5, fontface = "bold") +
  
  scale_x_continuous(breaks = seq(0, 15, by = 1)) +
  scale_y_continuous(labels = comma, expand = c(0, 0)) +
  labs(title = "Distance to Nearest School",
       subtitle = "Green zone indicates reasonable walking distance",
       x = "Distance (km)", 
       y = "Frequency (Grid Cells)")




```

## Data : Housing Price Transformation
The raw price distribution (left) is highly **right-skewed**, violating the normality assumption required for OLS regression. By applying a **log-transformation** (right), the data approximates a normal distribution, making it suitable for our later analysis model.

```{r}
#| layout-ncol: 2
#| echo: false

print(p1_price_raw)
print(p2_price_log)
```


## Data : School Quality & Proximity
These two variables directly address our research question regarding the impact of school quality and distance on housing prices.

```{r}
#| layout-ncol: 2
#| echo: false

print(p6_social)
print(p7_distance)
```

## Data : Housing Characteristics
Our sample captures the diversity of the housing stock in NRW. We observe a bimodal age structure (Post-war reconstruction vs. Modern development) and a log-normally distributed plot size (Median: 531 sqm).
```{r}
#| layout-ncol: 2
#| echo: false

print(p3_age)
print(p4_area_log)
```

## Data Validation

::: {.columns}

::: {.column width="40%"}
       
\textbf{Sanity Check Analysis}

- A clear *positive correlation* is observed.
- This confirms economic intuition: *larger houses cost more*.
- The plot helps visually identify *extreme values* for further cleaning.
:::

::: {.column width="60%"}
```{r}
#| out-width: "100%"
#| fig-align: "center"
#| echo: false

print(p5_sanity)
```
:::

:::
  
## Empirical approaches in the literature

•⁠  ⁠In the paper from Black & Machin, they identified six main approaches to analyze the relationship between school characteristics and housing prices:

1.⁠ ⁠Regression based estimates

2.⁠ ⁠Parametric and non-parametric modelling of unobservable factors

3.⁠ ⁠Instrumental variables approaches

4.⁠ ⁠Discontinuity methods using administrative boundaries

5.⁠ ⁠Difference-in-difference, repeat sales and quasi-experimental methods

6.⁠ ⁠Combined methods


---

## Our Method

•⁠  ⁠Hedonic Pricing Model: Apply a linear regression model to analyze the impact of the social index of schools on housing prices

•⁠  ⁠Proximity Effect: Incorporate distance to schools as an additional variable to examine whether proximity strengthens or weakens the effect of the social index on housing prices

•⁠  ⁠Spatial Model: Utilize a Spatial Lag Model or Spatial Durbin Model to account for spatial dependencies and the influence of neighboring schools and properties on housing prices

•⁠  ⁠Interaction Effects: Investigate the interaction between the social index and distance to schools to assess how proximity modulates the effect of the social index on housing prices

## Shortcomings of our strategy

-   A regression-based hedonic model in the housing context might be prone to certain endogeneity problems due to omitted variables.
-   A reason for that is the extensive scope of relevant housing characteristics and demographic features

\vspace{0.5cm}

	⁠Source: @Machin_2011


## Our Next Steps 

•⁠  ⁠Finalize Data Cleaning: Ensure all data is clean, standardized, and ready for analysis.

•⁠  ⁠Model Application: Apply appropriate regression and spatial models to examine the impact of school quality and proximity on housing prices.

•⁠  ⁠Result Interpretation: Analyze the model outputs, interpret key findings, and identify trends.


•⁠  ⁠Prepare Report & Presentation: Summarize results, create visualizations, and prepare findings for presentation to stakeholders.

