---
title: "Distance to Schools and House Prices"
subtitle: "Hypotheses & Regression Strategy"
author: "Your Name"
format:
  beamer:
    theme: Madrid
    colortheme: default
    slide-level: 2
    aspectratio: 169
    keep-tex: false
    pdf-engine: xelatex
    include-in-header:
      text: |
        \usepackage{booktabs}
        \usepackage{graphicx}
        \usepackage{float}
        \usepackage{amsmath}
        \usepackage{bm}
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false
source("Code_Final.R")
```

## Distance to Schools and House Prices: Hypotheses & Regression Strategy

::: block
**H1: House prices are higher closer to schools** *(negative coefficients on distance variables)*
:::

\vspace{0.1cm}
\hrule \vspace{0.1cm}

**Empirical framework**

-   **Outcome:** $\ln(\text{Price})$

-   **Key variables:** distance to nearest **primary** school; distance to nearest **secondary** school

-   **Strategy:** estimate **continuous**, **quadratic**, and **binned** distance gradients

\vspace{0.1cm}

::::::: columns
:::: {.column width="48%"}
::: exampleblock
### Baseline OLS controls

*Standard housing attributes*

-   Living area

-   Plot area

-   Number of rooms

-   House age
:::
::::

:::: {.column width="48%"}
::: exampleblock
### Distance designs

*Alternative specifications*

-   **Continuous:**

    -   include quadratic term ($km^2$)

-   **Binned**

    -   0–3 km, 3–6 km, 6–9 km, \> 9 km
:::
::::
:::::::

## Regression Specifications {.smaller}

\footnotesize

::::::::: columns
::::: {.column width="49%"}
::: block
### 1. Naive (Unconditional)

$$\ln(P) = \alpha + \beta_1 D^{prim} + \beta_2 D^{sec} + \varepsilon$$

-   **Goal:** Establish raw spatial correlation.

-   **Risk:** High Omitted Variable Bias.
:::

\vspace{0.2cm}

::: exampleblock
### 2. Baseline (Preferred)

$$\ln(P) = \alpha + \mathbf{\beta D} + \mathbf{\gamma X} + \varepsilon$$

-   **Controls (**$\mathbf{X}$): Area, Plot, Rooms, Age.

-   **Hypothesis:** $\beta < 0$ (Negative semi-elasticity).
:::
:::::

::::: {.column width="49%"}
::: block
### 3. Non-Linearity (Polynomial)

$$\ln(P) = \text{Base} \dots + \delta (D)^2 + \dots$$

-   **Logic:** Test for convex/concave decay.

-   **Test:** $\delta \neq 0$ implies varying MWTP.
:::

\vspace{0.2cm}

::: block
### 4&5. Robustness (Single-Type)

$$\ln(P) = \alpha + \beta_{k} D^{k} + \mathbf{\gamma X} + \varepsilon$$

-   **Method:** Estimate Primary/Secondary separately.

-   **Check:** Stability against multicollinearity.
:::
:::::
:::::::::

## Descriptive Evidence: Price-Distance Gradient

\vspace{-0.5cm}

\footnotesize

:::::: columns
::: {.column width="32%"}
**Negative Slope (H1)**

Clear decay in prices as distance increases across specifications.
:::

::: {.column width="32%"}
**Non-Linearity**

Steeper drop in the first 3-6 km, flattening afterwards.
:::

::: {.column width="32%"}
**Heterogeneity**

Primary schools show a visibly steeper gradient.
:::
::::::

```{r}
#| echo: false
#| fig-width: 14
#| fig-height: 5
#| fig-align: center
#| warning: false
#| message: false

library(ggplot2)
library(patchwork)

p_left <- fig_scatter_combined +
  labs(title = "Fig 1. Continuous Trend (Raw Scatter)") +
  theme(plot.title = element_text(size = 14, face = "bold"))

p_right <- fig_price_gradient +
  labs(title = "Fig 2. Discrete Gradient (Binned Means)") +
  theme(plot.title = element_text(size = 14, face = "bold"),
        axis.title.y = element_blank())

p_combined <- p_left + p_right +
  plot_layout(guides = "collect") & 
  theme(legend.position = "top",
        legend.box.margin = margin(0, 0, 0, 0),
        legend.margin = margin(0, 0, 0, 0))

print(p_combined)
```

## Main Results: Continuous Model {.smaller}

```{r}
#| echo: false
#| results: asis

modelsummary(
  models_cont,
  vcov      = "HC1",
  coef_map  = coef_map_cont,
  statistic = NULL,
  stars     = c("*" = .1, "**" = .05, "***" = .01),
  gof_map   = c("nobs", "r.squared", "adj.r.squared"),
  fmt       = 3,
  output    = "kableExtra"
) |>
  kableExtra::kable_styling(font_size = 7, latex_options = c("scale_down", "hold_position"))
```

## Main Results: Binned Model {.smaller}

```{r}
#| echo: false
#| results: asis

modelsummary(
  models_bin,
  vcov      = "HC1",
  statistic = NULL,
  stars     = c("*" = .1, "**" = .05, "***" = .01),
  gof_map   = c("nobs", "r.squared", "adj.r.squared"),
  fmt       = 3,
  output    = "kableExtra"
) |>
  kableExtra::kable_styling(font_size = 7, latex_options = c("scale_down", "hold_position"))
```

## Compare: Continuous vs Binned {.smaller}

```{r}
#| echo: false
#| results: asis

modelsummary(
  models_main_compare,
  vcov      = "HC1",
  coef_map  = coef_map_compare,
  coef_omit = "log_area|log_plot_area|zimmeranzahl|house_age|\\(Intercept\\)",
  statistic = NULL,
  stars     = c("*" = .1, "**" = .05, "***" = .01),
  gof_map   = c("nobs", "r.squared", "adj.r.squared"),
  fmt       = 3,
  output    = "kableExtra"
) |>
  kableExtra::kable_styling(font_size = 7, latex_options = c("scale_down", "hold_position")) |>
  kableExtra::footnote(general = "Binned coefficients are relative to the reference group: 0-3 km. Controls included but omitted from display.")
```

## Model comparison & robustness (main models)

Continuous fits slightly better;

binned is slightly more robust to influential points.

### Model comparison

```{r}
#| echo: false

tab_compare_all
```

## Model comparison & robustness (main models)

Continuous fits slightly better;

binned is slightly more robust to influential points.

### Robustness checks

```{r}
#| echo: false

tab_robust_checks
```

## Economic Mechanisms & Interpretations

\small

:::::: columns
::: {.column width="48%"}

::: block
### 1. Capitalization
Why are houses closer to schools more expensive?

- **Commuting Costs:** Time saved is directly capitalized into property values
- **Residential Sorting:** Families cluster here, increasing local demand
:::

\vspace{0.15cm}

::: exampleblock
### 2. Primary vs Secondary
Why is the effect stronger for primary schools?

- **Age Constraints:** Young children need escorting
- **Community Ties:** Primary schools define local neighborhoods
:::

:::

::: {.column width="48%"}

::: block
### 3. Non-Linearity
Why does the gradient flatten beyond 3km?

- **Walkability Premium:** Only exists within <3km distance
- **Substitution:** Public transit replaces driving at longer distances
:::

\vspace{0.15cm}

::: alertblock
### 4. Limitation
What is missing from this analysis?

- **Distance $\neq$ Quality:** Proximity captures accessibility, not desirability
- **Next Step:** Incorporate **Social Index** for school quality
:::

:::
::::::
